Try to use a time series analysis on one of the parks object

1. choose a park

2. make parks_long

3. parks_long has a Value column for each month of the year ranging from 1979 to 2023

4. make that column a time series object in the range with frequency 12 (monthly data)

load the library
```{r}
library(forecast)
library(shiny)
library(tidyverse)
library(plotly)
library(lubridate) #for graphic
library(ggplot2) # for make_date()
library(dplyr)
library(purrr)
library(sf)
library(leaflet)
```
read parks data

```{r}

parks <- read.csv("../data/park_visitor_data.csv")
parks <- parks|>
  mutate(across(-PARKCODE, ~as.integer(gsub(",", "", .))))

parks_long_ABLI <- parks |>
  filter(PARKCODE=='ABLI')|>
  select(-X,-PARKCODE, -TOTAL)|>
  gather(key = "Month", value = "Value", -YEAR)

#create a map to make Month column numeric
month_mapping <- c(JAN = 1, FEB = 2, MAR = 3, APR = 4, MAY = 5, JUN = 6,
                   JUL = 7, AUG = 8, SEP = 9, OCT = 10, NOV = 11, DEC = 12)

parks_long_ABLI <- parks_long_ABLI %>%
  mutate(MonthNumeric = month_mapping[Month])

# Create a new 'Date' column
parks_long_ABLI <- parks_long_ABLI %>%
  mutate(Date = make_date(YEAR, MonthNumeric))|>
  arrange(Date)
```
Save this parks_long dataframe as binary object

```{r}
parks_long_ABLI <- saveRDS(parks_long_ABLI, '../data/parks_long_ABLI.rds')
```


Now let's read this file

```{r}
parks_long_ABLI_try <- readRDS('../data/parks_long_ABLI.rds')
```


Create a time series object based on the Value column in the dataframe

```{r}
visitation_ABLI <- ts(
  parks_long_ABLI_try$Value, 
  start = c(1979, 1),
  end = c(2023, 12),
  frequency = 12)
```

plot the time series object 

```{r}
plot(visitation_ABLI)
```
```{r}



# Plot autocorrelation function with larger labels and fonts
par(cex.lab = 1.5, cex.axis = 1.5, cex.main = 1.5, cex.sub = 1.5)
acf_result <- acf(visitation_ABLI)
```
Fit a simple model

visitation_ABLI(t) = Constant + e_t

e_t is noise 
assumption : the noise is normally distributed ~ N(0,sigma)

```{r}

avg_model <- arima(visitation_ABLI, c(0,0,0))
```

look at the structure of the model 

```{r}
str(avg_model)
```
Forecast 
```{r}
avg_forecast <- forecast(avg_model, 48)
```


take a look at the forecast

```{r}
str(avg_forecast)
```
look at the mean value predicted 

```{r}
avg_forecast$mean
```

No variation in the predicted value as the model did not have any such time dependence
```{r}
plot(visitation_ABLI)
lines(avg_forecast$mean,col = "red")
```
Do this with tidyverse and ggplot ??

Let's add the predicted values to the dataframe

```{r}
parks_long_ABLI_try$predicted_values <- data.frame(datetime_column = x, y_values = y, arima_predictions = avg_forecast$mean)


```


```{r}
# Plotting the original data
ggplot(parks_long_ABLI_try, aes(x = Date, y = Value)) +
  geom_line(color = "blue")
```


```{r}
autoplot(avg_forecast)
```


Non-seasonal ARIMA model 

```{r}
arima_model <- auto.arima(visitation_ABLI, seasonal = FALSE)
arima_forecast <- forecast(arima_model, 48, level= c(50,95))
autoplot(arima_forecast)
```


Seasonal ARIMA model

```{r}
seasonal_arima_model <- auto.arima(visitation_ABLI)
seasonal_arima_forecast <- forecast(seasonal_arima_model, 48, level= c(50,95))
autoplot(seasonal_arima_forecast)
```
